Report on Late Winter2013 Production: Image Differencing

### 2013-03-26 W13report.tex
Filling out early section on goals, work performed.

We're done!

@@

The focus of this production is to optimize and quantify the false
positive rate in the Psf--matched subtraction of a science image and a
reference image (referred to as the template image below).  The metric
we optimize is the number of false detections that are detected and
measured on the image difference.  We calculate several additional
quality metrics during the image subtraction process, and correlate
them with the number of false detections, as a means to predict the
number of false detections we may expect downstream from the image
subtraction.

@@

We assume that science image $S(x,y)$ can be modeled as a convolution
of the template image $T(x,y)$ by a PSF--matching kernel $K(u,v,x,y)$.
We further assume that the kernel may be decomposed using a set of
basis functions $K(u,v) = \sum_i a_i K_i(u,v)$, where the coefficients
in front of each basis are determined through:
%
\begin{eqnarray}
C_i & \equiv & (K_i \otimes T) \\ \nonumber
b_{i}  & = & \sum_{x,y} {{C_i(x,y) S(x,y)}\over{\sigma^2(x,y)}}   \nonumber \\ 
M_{ij} & = & \sum_{x,y} {{C_i(x,y) C_j(x,y)}\over{\sigma^2(x,y)}}  \nonumber \\ 
a_{i}  & = & M^{-1}_{ij} b_{j} \nonumber 
\label{eq-soln}
\end{eqnarray}
where $\sigma^2(x,y)$ is the per--pixel variance stored in the {\tt
  variance} plane of each LSST {\tt exposure}.  To generate a
spatially varying model for the kernel, we assume that the relative
weights of the basis functions $a_i$ themselves vary spatially,
i.e. $K(u,v,x,y) = \sum_i a_i(x,y) K_i(u,v)$.  We also assume that
there is a differential background between the two images $b(x,y)$
that may be fit for using a low--order Chebyshev polynomial.  The
image difference is then calculated through $D(x,y) = S(x,y) - T(x,y)
\otimes K(u,v,x,y) - b(x,y)$.

The basis functions themselves $K_i(u,v)$ are a degree of freedom in
this problem.  The original implementations \citep{Alard98,Alard00}
used a set of 3 Gaussians (represented by our {\tt Config} variable
nGauss), each with a different sigma (sigGauss), and each modified by
a Laguerre polynomial to a given order (degGauss).  Subsequent studies
\citep[e.g.][]{2007AN....328...16I} have suggested that a constant
ratio be maintained between the different Gaussian widths, such that
$\sigma_{i+1} = \beta \times \sigma_{i}$.  We use the value $\beta =
2.0$ for this production.  We set the overall scale for sigma by
noting that, under the assumption that the Psfs of the images are
Gaussian ($\sigma_S$ for the science image and $\sigma_T$ for the
template image), the sigma of the matching kernel should be simply
$\sigma_K^2 = \sigma_S^2 - \sigma_T^2$.  We use this width for the
central Gaussian in the basis sequences where we use more than one
Gaussian, or the sole Gaussian in the sequences when we use only one.

Detection on the difference image occurs through correlation of
$D(x,y)$ with the science image's Psf, yielding optimally filtered
detection image $D'(x,y) = D(x,y) \circ Psf_S(x,y,u,v)$ where $\circ$
denotes correlation.  The values of the pixels in $D'(x,y)$ provide a
maximum likelihood estimate of there being a point source at that
position.  Detection occurs by simply finding pixels that are more
than N sigma above the variance.

In this production, we investigated two orders of operations for the
optimal filtering by the Psf.  In the first method, called
prefiltering, we Psf--match the template image to a pre--filtered
science image:
\begin{eqnarray}
D_{Pre}(x,y) & = & S(x,y) \circ Psf_S(x,y,u,v) - T(x,y) \otimes K(u,v,x,y) \nonumber \\ 
D'_{Pre}(x,y) & = & D_{Pre}(x,y). \nonumber 
\end{eqnarray}
By pre--smoothing the image that we are matching the template to --
$S(x,y) \circ Psf_S(x,y,u,v)$ -- we more frequently avoid the case
of psf--matching {\it deconvolution}, where the FWHM of $T$ is
narrower than $S$.  This pre--filtering increases the FWHM of the
image by $\sqrt{2}$.  The Psf matching kernels will need to be
correspondingly larger to account for the larger Psfs.  We are able to
run detection directly on this image.

In the second, which is the classical postfiltering implementation of
the technique, we create a difference image that is then
post--filtered with the Psf for detection:
\begin{eqnarray}
D_{Post}(x,y) & = & S(x,y) - T(x,y) \otimes K(u,v,x,y) \nonumber \\ 
D'_{Post}(x,y) & = & D_{Post}(x,y) \circ Psf_S(x,y,u,v).  \nonumber 
\end{eqnarray}

@@

@@

Next, we use the DiaCatalogSourceSelector to query the ImSim reference
catalog for appropriate sources to use for Psf matching.  This
selector allows the user to specify the brightness and color range of
the objects, toggle star or galaxy selection, and to include variable
objects or not.  We select a fraction (1/5) of these objects to serve
as a control sample, in order to assess the effectiveness of
$K(u,v,x,y)$ to subtract objects that were {\it not} used in the least
squares fit.

@@

Images $S(x,y)$ (optionally pre-filtered with $Psf_S(x,y,u,v)$) and
$T(x,y)$ are sent to ImagePsfMatchTask, whose purpose is to fit for
$K(u,v,x,y)$ and to produce $D(x,y)$.  The first step is to use the
Psf Gaussian sigmas to define the {\it sizes} of the Gaussians shapes
in $K_i(u,v)$.  As mentioned above, the main width is chosen to be
$\sqrt{\sigma_S^2 - \sigma_T^2}$.  We investigate variations on the
configuration where we use one Gaussian with the above width, and
where we use three Gaussians, the central of which has this width and
the others are a factor of $\beta = 2.0$ smaller and larger.  We
modify each of the Gaussians by a set of Laguerre polynomials of a
given order.  The smallest Gaussian is modified to the highest order,
with the others modified by order//2.  Figure~\ref{basis} provides an
example basis set with this configuration.  The total number of shapes
in the basis are: $\sum_i^{\rm nGauss} ({\rm degGauss}_i+1)\times({\rm
  degGauss}_i+2)/2$ \footnote{ A notable exception to this algorithm
  is found in the case where $\sigma_T > \sigma_S$, i.e. a sharpening
  must be done to the template to match it to the science image.  This
  is the case of deconvolution.  A theoretical prescription for basis
  design in this case is given in \cite{0266-5611-26-8-085002}.  In
  this algorithm, nGauss is {\it always} fixed to value 3, and
  degGauss is always fixed to 3.  The widths of the Gaussians are
  determined through the algorithm specified in
  \cite{0266-5611-26-8-085002}, using as inputs the sequence of
  Gaussians that we would have used to match a Gaussian of width
  $\sigma_S$ to $\sigma_T$ (i.e. as if we would have convolved the
  science image and not the template image).  There are a variety of
  reasons that actually operating on the science image is undesirable,
  including: the science image is lower S/N than the template; the
  science image is likely to have masked defects, including cosmic
  rays and bad pixels, whose reach would be expanded due to the
  convolution; unmasked features such as faint cosmic rays are more
  likely to look Psf--like after convolution (cosmic rays will likely
  have been sigma clipped out of the coadd).  However, it is not clear
  that deconvolving the template image is in principle any better.  A
  far better option, which we explore here, is to pre--filter the
  science image, which will in the vast majority cases provide a
  larger Psf (larger by $\sqrt{2}$) to match the template to.  This is
  the case for v6866601 in this study, where the postfiltered analysis
  yields deconvolution of the template, while the prefiltered analysis
  does not.  }.
%
The dimensions of the Psf matching kernel are
chosen to be 6 times the largest Gaussian width, with a minimum
kernelSize of 21x21 pixels.  For kernels that are significantly
smaller than this, the Gaussians have significant (non--zero) power at
the kernel boundaries, leading to square systematic artifacts at the
scale of the kernel in the difference images.

@@

We next run the SourceDetectionTask on the difference image, using a
``polarity'' of {\it both} to search for both positive and negative
deviations.  We use a nominal detection threshold of 5 sigma.  In the
case of pre--filtering, we detect on the difference image directly.
In post--filtering, we correlate the difference image with its Psf (by
design, the same as the Psf of $S$) before detection is run.  All else
being equal, the images immediately preceding the detection step
($D'$) should be {\it exactly} the same.  However, differences in the
shapes and sizes of the Psf--matching kernels used in the two cases
may cause systematic differences in the images.  Detections are merged
using a grow radius of 2 pixels.

Finally, we run DipoleMeasurementTask on the detections to create {\tt
  DiaSources}.  DipoleMeasurementTask is an extension of
SourceMeasurementTask that includes specific dipole measurement
algorithms.  Measurement is necessarily different in the pre--filtered
and post--filtered data.  In particular, in post--filtering, detection
happens on the filtered maximum likelihood image and source
measurement on the associated pixels in the unfiltered difference
image.  In pre--filtering, the latter does not exist, and thus
detection and measurement happen on the same image.  This restricts
the types of measurements that may be made on the sources.  For
example, the Psf flux is merely the fitted amplitude of the peak in
the image, and the Psf centroid its center.  While these measurements
are trivial to implement, additional work needs to be done on how to
implement the remainder of the measurement suite in a manner
comparable to the measurements on post--filtered data.

@@

We have examined the distribution of these false detections across the
image, and find no significant overdensities near the borders of the
images, which would be expected in cases of under or over--fitting of
the various spatial models (astrometry, backgrounds, Psf--matching
kernels).  Figure~\ref{edgedist} provides the distribution of
distances from the edge of the sensor for all false detections in {\it
  blue}, with a random sample of points in the image given in {\it
  green}.  Uncertainties are plotted using the square root of the
number of points in each bin, and the histograms are normalized to
provide a probability density.  The takeaway is that we find no
significant correlation of false detections with the boundaries of the
fitting functions.

@@

Figure~\ref{kernel_offsets} demonstrates the ability of the
Psf--matching kernel to redistribute flux to account for bulk
astrometric misalignments.  Since these offsets are common--mode
amongst all {\tt KernelCandidates}, the basis shapes themselves drive
the pipeline's ability to compensate for them.  Having a Gaussian
function whose sigma is commensurate with the amplitude of the offset
enables the software to place significant power in this basis term.  We
expect that, if these misalignments are a vector field, the ability of
the spatial kernel model to account for the spatial variation of the
misregistration will begin to drive the solution.  The take--away
from this analysis is that the Psf--matching basis appears flexible
enough to account for, individually, random scatter and bulk offsets
in the WCS solution.  This narrows the focus down to understanding how
spatial variation in the WCS may drive the difference image quality,
and what requirements need to be put both on the kernel basis and its
spatial variation.

@@

Because these images will have gone through multiple convolutions
before reaching the detection stage, it is expected that the
propagated per-pixel variance will not provide an exact representation
of the empirical variance in the image planes \citep{Price-Stacking}.
Because the astrometric resampling uses a warping kernel that is
designed to preserve the noise properties of the images, we expect
that the kernel convolution and Psf--filtering will be the largest
sources of error.

To quantify these errors, we first compared the empirical variance in
the image planes with the level tracked by the variance planes.  We
calculated the ratio of the interquartile range of unmasked pixels in
the image plane, multiplied by 0.741 and then squared, with the median
of unmasked pixels from the variance plane.  The former represents the
empirical variance, while the latter the propagated variance.  We
investigate their ratio (empirical/propagated) at 2 stages in the
processing: after creation of the difference image (image $D$), and
immediately preceding detection (image $D'$).  Recall that for
prefiltering, $D = D'$.  These results are summarized in
Table~\ref{tab-variance1}.  For prefiltering, we find that the
variance plane typically underestimates the true variance by $1\%$ for
all visits.  When using postfiltering, the deconvolution visit
v6866601 yields a similar underestimate in the difference image, but
for the other two visits the variance is tracked to better than 1\%,
with a small RMS.  However, when postfiltering with the Psf, the
variance becomes {\it overestimated} by nearly a factor of two for
deconvolution visit v6866601, and {\it underestimated} by $4-5\%$ for
the other visits.

We further investigate the variance properties of the images that are
input to the detection stage, i.e. the difference image itself in the
case of prefiltering, and the difference image correlated with its Psf
in postfiltering.  These results are summarized in
Table~\ref{tab-variance2}.  We find that, empirically, the images have
very similar noise properties at the detection stage, with the
deconvolution visit having larger variance by $1\%$ but the other
visits having essentially equal noise properties, with a bias towards
the postfiltered data having slightly higher variance (while the RMS
is small, all values are greater than 1.0).  The medians of the
variance planes tell a different story though.  For the deconvolution
visit, the median per--pixel variance is $70\%$ higher.  While this
will suppress the detections of false positives, it will also
significantly suppress the detection efficiency for true variability.
For the other visits, the prefiltered per--pixel variance is
relatively {\it higher} by $3-4\%$.  This will decrease the
sensitivity of the prefiltered data to false positives; as shown
above, the rate is indeed smaller using the prefiltering pipeline.
This analysis demonstrates that this is primarily due to an {\it
  underestimate} of the per--pixel variance in the postfilter
processing: while the prefiltered variance is a 1\% underestimate, the
postfiltered variance is a 5\% underestimate.  As outlined by
\cite{Price-Stacking}, this may be due to differences in pixel
covariance when undergoing multiple convolutions, which are not
currently tracked.  We have not yet definitively demonstrated that
this is the case here.

@@

For this production, the understanding of the algorithm and
optimization of the false detection rates took priority over
optimizing the computational performance of the algorithms.  In
particular, debugging hooks were impolemented and workarounds of
problems were devised with the focus on understanding the core
Psf--matching algorithm.  Therefore there are several key ways in
which the production as--implemented will differ from how the code is
used in Nightly and Data Release Processing.  We outline these
differences below.

@@

The reported {\tt imageDifference:detection} task performance is an
accurate representation of its computational requirements.  However,
the {\tt imageDifference:dipolemeasurement} task represents a
significant overestimate of the true dipole measurement requirements.
In particular, the Psf--flux dipole measurement is explicitly
sub--optimal, in that its centroid fitting algorithm is a 4--deep {\tt
  for} loop.

@@

\item When convolving to Psf--match, variance planes of postfiltered
  detection images have {\it lower} variance than their prefiltered
  counterparts by $2-3\%$ (Table~\ref{tab-variance2}).  These images
  have gone through one additional convolution compared to the
  prefiltered data.

@@

\item The numbers of false detections that associate with Sources is
  shown to correlate with the amount of smoothing that occurs in the
  Psf matching, in the sense that more smoothing leads to fewer false
  detections (Figure~\ref{}).  In this sense, there is a strong
  preference for running in prefiltered mode.

@@

\begin{figure}
\begin{center}
\includegraphics[width=0.9\textwidth]{figures/normal2.eps} \\
\includegraphics[width=0.9\textwidth]{figures/preconv2.eps} \\
\end{center}
\caption{Pairs of LOCAL,SPATIAL difference images in the case of
  successful image subtraction, similar to
  Figure~\ref{fig_galleryBad}.  The {\it top} set of data are from
  postfiltered analysis, while the {\it lower} set are from
  prefiltered data.  Note that the noise is smoother in the
  prefiltered data due to correlation with the Psf. }
\label{fig_galleryGood}
\end{figure}

@@

\begin{figure}
\includegraphics[width=0.8\textwidth]{figures/shift.eps} \\
\caption{Figure illustrating the ability of the kernel basis to
  compensate, locally, for the effects of misregistration.  Each
  element of the grid shows a single {\tt KernelCandidate} local
  difference image (visit=6866603, doPreFilter=True).  Columns
  indicate the Gaussian sigma of the basis, which is designed to have
  one Gaussian with the prescribed width, modified to fourth order.
  Rows indicate the amplitude of the astrometric offset that was
  introduced into the Source positions before image-to-image
  registration.  In the {\it top} row, we see that even with a small
  astrometric offset, the sigma=2 basis is unable to produce a quality
  subtraction, because the width is inappropriate for matching the two
  input Psfs (the theoretical Gaussian sigma here is 3.4 pixels).  At
  sigma=4 the basis is able to produce a quality subtraction, and by
  sigma=5 the basis is too large to match the Psfs.  In {\it row 4},
  we show that even for an astrometric offset of 4 pixels (0.8''), the
  sigma=4 basis can produce a reasonable difference image without
  structured residuals.  However, by {\it row 5}, with an offset of 6
  pixels, none of the Gaussians provide a successful local difference
  image.  The ability of the basis to compensate for bulk astrometric
  offsets is a function of the kernel width, and the kernel width is
  itself a function of the two input Psf FWHMs.  Thus there is a
  complicated dependence of the ability of the kernel to shift the
  centroids of stars.  Roughly, in our implementation, the basis may
  compensate for local astrometric residuals at a scale up to $\beta
  \times \sqrt{\sigma_S^2 - \sigma_T^2}$, which is the width of the
  largest Gaussian in our bases.}
\label{kernel_offsets}
\end{figure}

>>>>

Summary.

@@

The focus of this production was to optimize and quantify the false
detection rate in the PSF--matched subtraction of a science image and
a reference image (referred to as the template image below).  We used
simulated images generated using ideal conditions, and with no
intrinsic variability in the sources.  The metric we optimized is the
number of false detections that are detected and measured on the image
difference.  As the input were simulated with no variability, any
detections found in the difference images are ``false positives''.  We
calculated several additional quality metrics during the image
subtraction process, and correlated them with the number of false
detections, as a means to predict the number of false detections we
may expect downstream from the image subtraction.  Once we found an
optimum configuration, we perturbed the solution along multiple
dimensions to understand how the false detection rate responded.

@@

We outline the image subtraction process below.  We assume that science
image $S(x,y)$ can be modeled as a convolution of the template image
$T(x,y)$ by a PSF--matching kernel $K(u,v,x,y)$.  The images will have
different point spread functions (PSFs), which are the time--averaged
transfer functions of a point source through the Earth's atmosphere,
telescope optics, and into the silicon of the detector before being
read out.  The essence of image subtraction is to match the PSFs of
these two images so that they may be subtracted pixel by pixel.  We
further assume that the PSF--matching kernel may be decomposed using a
set of basis functions $K(u,v) = \sum_i a_i K_i(u,v)$, where the
coefficients in front of each basis are determined through:
%
\begin{eqnarray}
C_i & \equiv & (K_i \otimes T) \\ \nonumber
b_{i}  & = & \sum_{x,y} {{C_i(x,y) S(x,y)}\over{\sigma^2(x,y)}}   \nonumber \\ 
M_{ij} & = & \sum_{x,y} {{C_i(x,y) C_j(x,y)}\over{\sigma^2(x,y)}}  \nonumber \\ 
a_{i}  & = & M^{-1}_{ij} b_{j} \nonumber 
\label{eq-soln}
\end{eqnarray}
where $\sigma^2(x,y)$ is the per--pixel variance stored in the {\tt
  variance} plane of each LSST {\tt exposure}.  To generate a
spatially varying model for the kernel, we assume that the relative
weights of the basis functions $a_i$ themselves vary spatially,
i.e. $K(u,v,x,y) = \sum_i a_i(x,y) K_i(u,v)$.  We also assume that
there is a differential background between the two images $b(x,y)$
that may be fit for using a low--order Chebyshev polynomial.  The
image difference is then calculated through $D(x,y) = S(x,y) - T(x,y)
\otimes K(u,v,x,y) - b(x,y)$.

The basis functions themselves $K_i(u,v)$ are a degree of freedom in
this problem.  The original implementations \citep{Alard98,Alard00}
used a set of 3 Gaussians (represented by our {\tt Config} variable
nGauss), each with a different sigma (sigGauss), and each modified by
a Laguerre polynomial to a given order (degGauss).  Subsequent studies
\citep[e.g.][]{2007AN....328...16I} have suggested that a constant
ratio be maintained between the different Gaussian widths, such that
$\sigma_{i+1} = \beta \times \sigma_{i}$.  We use the value $\beta =
2.0$ for this production.  We set the overall scale for sigma by
noting that, under the assumption that the PSFs of the images are
Gaussian ($\sigma_S$ for the science image and $\sigma_T$ for the
template image), the sigma of the matching kernel should be simply
$\sigma_K^2 = \sigma_S^2 - \sigma_T^2$.  We use this width for the
central Gaussian in the basis sequences where we use more than one
Gaussian, or the sole Gaussian in the sequences when we use only one.

Detection on the difference image occurs through correlation of
$D(x,y)$ with the science image's PSF, yielding optimally filtered
detection image $D'(x,y) = D(x,y) \circ PSF_S(x,y,u,v)$ where $\circ$
denotes correlation.  The values of the pixels in $D'(x,y)$ provide a
maximum likelihood estimate of there being a point source at that
position.  Detection occurs by simply finding pixels that are more
than N sigma above the variance.

In this production, we investigated two orders of operations for the
optimal filtering by the PSF.  In the first method, called
prefiltering, we PSF--match the template image to a pre--filtered
science image:
\begin{eqnarray}
D_{Pre}(x,y) & = & S(x,y) \circ PSF_S(x,y,u,v) - T(x,y) \otimes K(u,v,x,y) \nonumber \\ 
D'_{Pre}(x,y) & = & D_{Pre}(x,y). \nonumber 
\end{eqnarray}
By pre--smoothing the image that we are matching the template to --
$S(x,y) \circ PSF_S(x,y,u,v)$ -- we more frequently avoid the case
of PSF--matching {\it deconvolution}, where the FWHM of $T$ is
narrower than $S$.  This pre--filtering increases the FWHM of the
image by $\sqrt{2}$.  The PSF matching kernels will need to be
correspondingly larger to account for the larger PSFs.  We are able to
run detection directly on this image.

In the second, which is the classical postfiltering implementation of
the technique, we create a difference image that is then
post--filtered with the PSF for detection:
\begin{eqnarray}
D_{Post}(x,y) & = & S(x,y) - T(x,y) \otimes K(u,v,x,y) \nonumber \\ 
D'_{Post}(x,y) & = & D_{Post}(x,y) \circ PSF_S(x,y,u,v).  \nonumber 
\end{eqnarray}

@@

Tasks that were completed in the run--up to production include:
implementing prefiltering for image subtraction (#2551); implementing
source selection for image subtraction (#2554, #2596); implementing
measurement for difference image source (DiaSource) characterization
(#2595); implementing RegisterTask for image--to--image registration
(#2562, #2650); and defining and implementing metrics for testing
ip\_diffim (#2636).  We describe in detail the overall pipeline that
was established as the results of these tasks below.

@@

Next, we use the DiaCatalogSourceSelector to query the ImSim reference
catalog for appropriate sources to use for PSF matching.  This
selector allows the user to specify the brightness and color range of
the objects, toggle star or galaxy selection, and to include variable
objects or not.  We select a fraction (1/5) of these objects to serve
as a control sample, in order to assess the effectiveness of
$K(u,v,x,y)$ to subtract objects that were {\it not} used in the least
squares fit.

@@

Images $S(x,y)$ (optionally pre-filtered with $PSF_S(x,y,u,v)$) and
$T(x,y)$ are sent to ImagePsfMatchTask, whose purpose is to fit for
$K(u,v,x,y)$ and to produce $D(x,y)$.  The first step is to use the
PSF Gaussian sigmas to define the {\it sizes} of the Gaussians shapes
in $K_i(u,v)$.  As mentioned above, the main width is chosen to be
$\sqrt{\sigma_S^2 - \sigma_T^2}$.  We investigate variations on the
configuration where we use one Gaussian with the above width, and
where we use three Gaussians, the central of which has this width and
the others are a factor of $\beta = 2.0$ smaller and larger.  We
modify each of the Gaussians by a set of Laguerre polynomials of a
given order.  The smallest Gaussian is modified to the highest order,
with the others modified by order//2.  Figure~\ref{basis} provides an
example basis set with this configuration.  The total number of shapes
in the basis are: $\sum_i^{\rm nGauss} ({\rm degGauss}_i+1)\times({\rm
  degGauss}_i+2)/2$ \footnote{ A notable exception to this algorithm
  is found in the case where $\sigma_T > \sigma_S$, i.e. a sharpening
  must be done to the template to match it to the science image.  This
  is the case of deconvolution.  A theoretical prescription for basis
  design in this case is given in \cite{0266-5611-26-8-085002}.  In
  this algorithm, nGauss is {\it always} fixed to value 3, and
  degGauss is always fixed to 3.  The widths of the Gaussians are
  determined through the algorithm specified in
  \cite{0266-5611-26-8-085002}, using as inputs the sequence of
  Gaussians that we would have used to match a Gaussian of width
  $\sigma_S$ to $\sigma_T$ (i.e. as if we would have convolved the
  science image and not the template image).  There are a variety of
  reasons that actually operating on the science image is undesirable,
  including: the science image is lower S/N than the template; the
  science image is likely to have masked defects, including cosmic
  rays and bad pixels, whose reach would be expanded due to the
  convolution; unmasked features such as faint cosmic rays are more
  likely to look PSF--like after convolution (cosmic rays will likely
  have been sigma clipped out of the coadd).  However, it is not clear
  that deconvolving the template image is in principle any better.  A
  far better option, which we explore here, is to pre--filter the
  science image, which will in the vast majority cases provide a
  larger PSF (larger by $\sqrt{2}$) to match the template to.  This is
  the case for v6866601 in this study, where the postfiltered analysis
  yields deconvolution of the template, while the prefiltered analysis
  does not.  }.
%
The dimensions of the PSF matching kernel are
chosen to be 6 times the largest Gaussian width, with a minimum
kernelSize of 21x21 pixels.  For kernels that are significantly
smaller than this, the Gaussians have significant (non--zero) power at
the kernel boundaries, leading to square systematic artifacts at the
scale of the kernel in the difference images.

@@

We next run the SourceDetectionTask on the difference image, using a
``polarity'' of {\it both} to search for both positive and negative
deviations.  We use a nominal detection threshold of 5 sigma.  In the
case of pre--filtering, we detect on the difference image directly.
In post--filtering, we correlate the difference image with its PSF (by
design, the same as the PSF of $S$) before detection is run.  All else
being equal, the images immediately preceding the detection step
($D'$) should be {\it exactly} the same.  However, differences in the
shapes and sizes of the PSF--matching kernels used in the two cases
may cause systematic differences in the images.  Detections are merged
using a grow radius of 2 pixels.

Finally, we run DipoleMeasurementTask on the detections to create {\tt
  DiaSources}.  DipoleMeasurementTask is an extension of
SourceMeasurementTask that includes specific dipole measurement
algorithms.  Measurement is necessarily different in the pre--filtered
and post--filtered data.  In particular, in post--filtering, detection
happens on the filtered maximum likelihood image and source
measurement on the associated pixels in the unfiltered difference
image.  In pre--filtering, the latter does not exist, and thus
detection and measurement happen on the same image.  This restricts
the types of measurements that may be made on the sources.  For
example, the PSF flux is merely the fitted amplitude of the peak in
the image, and the PSF centroid its center.  While these measurements
are trivial to implement, additional work needs to be done on how to
implement the remainder of the measurement suite in a manner
comparable to the measurements on post--filtered data.

@@

We have examined the distribution of these false detections across the
image, and find no significant overdensities near the borders of the
images, which would be expected in cases of under or over--fitting of
the various spatial models (astrometry, backgrounds, PSF--matching
kernels).  Figure~\ref{edgedist} provides the distribution of
distances from the edge of the sensor for all false detections in {\it
  blue}, with a random sample of points in the image given in {\it
  green}.  Uncertainties are plotted using the square root of the
number of points in each bin, and the histograms are normalized to
provide a probability density.  The takeaway is that we find no
significant correlation of false detections with the boundaries of the
fitting functions.

@@

Figure~\ref{kernel_offsets} demonstrates the ability of the
PSF--matching kernel to redistribute flux to account for bulk
astrometric misalignments.  Since these offsets are common--mode
amongst all {\tt KernelCandidates}, the basis shapes themselves drive
the pipeline's ability to compensate for them.  Having a Gaussian
function whose sigma is commensurate with the amplitude of the offset
enables the software to place significant power in this basis term.  We
expect that, if these misalignments are a vector field, the ability of
the spatial kernel model to account for the spatial variation of the
misregistration will begin to drive the solution.  The take--away
from this analysis is that the PSF--matching basis appears flexible
enough to account for, individually, random scatter and bulk offsets
in the WCS solution.  This narrows the focus down to understanding how
spatial variation in the WCS may drive the difference image quality,
and what requirements need to be put both on the kernel basis and its
spatial variation.

@@

Because these images will have gone through multiple convolutions
before reaching the detection stage, it is expected that the
propagated per-pixel variance will not provide an exact representation
of the empirical variance in the image planes \citep{Price-Stacking}.
Because the astrometric resampling uses a warping kernel that is
designed to preserve the noise properties of the images, we expect
that the kernel convolution and PSF--filtering will be the largest
sources of error.

To quantify these errors, we first compared the empirical variance in
the image planes with the level tracked by the variance planes.  We
calculated the ratio of the interquartile range of unmasked pixels in
the image plane, multiplied by 0.741 and then squared, with the median
of unmasked pixels from the variance plane.  The former represents the
empirical variance, while the latter the propagated variance.  We
investigate their ratio (empirical/propagated) at 2 stages in the
processing: after creation of the difference image (image $D$), and
immediately preceding detection (image $D'$).  Recall that for
prefiltering, $D = D'$.  These results are summarized in
Table~\ref{tab-variance1}.  For prefiltering, we find that the
variance plane typically underestimates the true variance by $1\%$ for
all visits.  When using postfiltering, the deconvolution visit
v6866601 yields a similar underestimate in the difference image, but
for the other two visits the variance is tracked to better than 1\%,
with a small RMS.  However, when postfiltering with the PSF, the
variance becomes {\it overestimated} by nearly a factor of two for
deconvolution visit v6866601, and {\it underestimated} by $4-5\%$ for
the other visits.

We further investigate the variance properties of the images that are
input to the detection stage, i.e. the difference image itself in the
case of prefiltering, and the difference image correlated with its PSF
in postfiltering.  These results are summarized in
Table~\ref{tab-variance2}.  We find that, empirically, the images have
very similar noise properties at the detection stage, with the
deconvolution visit having larger variance by $1\%$ but the other
visits having essentially equal noise properties, with a bias towards
the postfiltered data having slightly higher variance (while the RMS
is small, all values are greater than 1.0).  The medians of the
variance planes tell a different story though.  For the deconvolution
visit, the median per--pixel variance is $70\%$ higher.  While this
will suppress the detections of false positives, it will also
significantly suppress the detection efficiency for true variability.
For the other visits, the prefiltered per--pixel variance is
relatively {\it higher} by $3-4\%$.  This will decrease the
sensitivity of the prefiltered data to false positives; as shown
above, the rate is indeed smaller using the prefiltering pipeline.
This analysis demonstrates that this is primarily due to an {\it
  underestimate} of the per--pixel variance in the postfilter
processing: while the prefiltered variance is a 1\% underestimate, the
postfiltered variance is a 5\% underestimate.  As outlined by
\cite{Price-Stacking}, this may be due to differences in pixel
covariance when undergoing multiple convolutions, which are not
currently tracked.  We have not yet definitively demonstrated that
this is the case here.

@@

For this production, the understanding of the algorithm and
optimization of the false detection rates took priority over
optimizing the computational performance of the algorithms.  In
particular, debugging hooks were implemented and workarounds of
problems were devised with the focus on understanding the core
PSF--matching algorithm.  Therefore there are several key ways in
which the production as--implemented will differ from how the code is
used in Nightly and Data Release Processing.  We outline these
differences below.

@@

The reported {\tt imageDifference:detection} task performance is an
accurate representation of its computational requirements.  However,
the {\tt imageDifference:dipolemeasurement} task represents a
significant overestimate of the true dipole measurement requirements.
In particular, the PSF--flux dipole measurement is explicitly
sub--optimal, in that its centroid fitting algorithm is a 4--deep {\tt
  for} loop.

@@

\item When convolving to PSF--match, variance planes of postfiltered
  detection images have {\it lower} variance than their prefiltered
  counterparts by $2-3\%$ (Table~\ref{tab-variance2}).  These images
  have gone through one additional convolution compared to the
  prefiltered data.

@@

\item The numbers of false detections that associate with Sources is
  shown to correlate with the amount of smoothing that occurs in the
  PSF matching, in the sense that more smoothing leads to fewer false
  detections (Figure~\ref{}).  In this sense, there is a strong
  preference for running in prefiltered mode.

@@

\begin{figure}
\begin{center}
\includegraphics[width=0.9\textwidth]{figures/normal2.eps} \\
\includegraphics[width=0.9\textwidth]{figures/preconv2.eps} \\
\end{center}
\caption{Pairs of LOCAL,SPATIAL difference images in the case of
  successful image subtraction, similar to
  Figure~\ref{fig_galleryBad}.  The {\it top} set of data are from
  postfiltered analysis, while the {\it lower} set are from
  prefiltered data.  Note that the noise is smoother in the
  prefiltered data due to correlation with the PSF. }
\label{fig_galleryGood}
\end{figure}

@@

\begin{figure}
\includegraphics[width=0.8\textwidth]{figures/shift.eps} \\
\caption{Figure illustrating the ability of the kernel basis to
  compensate, locally, for the effects of misregistration.  Each
  element of the grid shows a single {\tt KernelCandidate} local
  difference image (visit=6866603, doPreFilter=True).  Columns
  indicate the Gaussian sigma of the basis, which is designed to have
  one Gaussian with the prescribed width, modified to fourth order.
  Rows indicate the amplitude of the astrometric offset that was
  introduced into the Source positions before image-to-image
  registration.  In the {\it top} row, we see that even with a small
  astrometric offset, the sigma=2 basis is unable to produce a quality
  subtraction, because the width is inappropriate for matching the two
  input PSFs (the theoretical Gaussian sigma here is 3.4 pixels).  At
  sigma=4 the basis is able to produce a quality subtraction, and by
  sigma=5 the basis is too large to match the PSFs.  In {\it row 4},
  we show that even for an astrometric offset of 4 pixels (0.8''), the
  sigma=4 basis can produce a reasonable difference image without
  structured residuals.  However, by {\it row 5}, with an offset of 6
  pixels, none of the Gaussians provide a successful local difference
  image.  The ability of the basis to compensate for bulk astrometric
  offsets is a function of the kernel width, and the kernel width is
  itself a function of the two input PSF FWHMs.  Thus there is a
  complicated dependence of the ability of the kernel to shift the
  centroids of stars.  Roughly, in our implementation, the basis may
  compensate for local astrometric residuals at a scale up to $\beta
  \times \sqrt{\sigma_S^2 - \sigma_T^2}$, which is the width of the
  largest Gaussian in our bases.}
\label{kernel_offsets}
\end{figure}
